White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions

### 2018-01-22 papers/wdmerger_II/paper.tex
Rework the problem setup section

%==========================================================================
% Problem Setup
%==========================================================================
\section{White Dwarf Collisions}
\label{sec:collisions}

We start with a description of the problem setup and results of a typical collision
at low resolution. These results set the stage for our numerical methodology. This enables
a concrete discussion about the methodology using the language of how it affects the
collision outcome. In this setting, a concrete discussion is more fruitful than an abstract one.

\subsection{Problem Setup}
\label{sec:problemsetup}

@@

Our primary metric for this test is the amount of $^{56}$Ni generated in the collision,
as this is the parameter most directly related to the observable quantities of interest
for Type Ia supernovae. We collect information at the end of every timestep about the
total amount of nickel on the grid (in solar masses), and we use the maximum value of
this nickel mass. Plotfiles of relevant quantities are created every 0.01 seconds.
The timesteps are limited to ensure that every multiple of this plotfile interval is
reached exactly, which means an effective maximum timestep of 0.01 seconds.

\subsection{Baseline Results}
\label{sec:baseline_results}

Now let us examine the results for a collision of two 0.64 $\msolar$ WDs. We use
128 zones in the radial dimension, corresponding to a spatial resolution of 312.5 km.
Timesteps are limited only by the hydrodynamic stability constraint (and the plotfile
interval constraint). Resolutions in the range of a few hundred km have been a common
choice in earlier papers on white dwarf mergers and collisions. No adaptive mesh
refinement is used. For this configuration, the collision time is approximately $t_{\rm c} = 4.6\ \text{s}$,
and the detonation time is $t_{\rm d} = 7.43\ \text{s}$. The peak density at the detonation time
is $2.3 \times 10^7\ \text{g\,/\,cm}^3$. The detonation occurs at the center of the domain
where the two WDs initially meet.

@@

At high resolution the collision time, $t_{\rm c} = 4.6\ \text{s}$, is similar to the
collision time at low resolution. However, the detonation time is around $t_{\rm d} = 6.62\ \text{s}$,
which is significantly earlier than at low resolution due to the high resolution simulations
being better able to resolve the point of contact between the WDs. The peak density crosses
$10^7\ \text{g\,/\,cm}^3$ significantly earlier, and this is what triggers the burning sufficient
to generate an earlier detonation. Approximately speaking, the later $t_{\rm d}$ is, the more
nickel is produced. As time passes, more white dwarf material slams into the center, compressing
and heating up the gas. Both effects are favorable to production of nickel. So, a later $t_{\rm d}$
implies that more material can be converted to nickel.

>>>>

%==========================================================================
% Problem Setup
%==========================================================================
\section{White Dwarf Collision Setup}
\label{sec:collisions}

@@

We collect information at the end of every timestep about the total amount of
all of the species on the grid (in solar masses), and we use the mass of $^{56}$Ni
generated as a metric for this test, as this is the parameter most directly related
to the observable quantities of interest for Type Ia supernovae. Since, as we will see
below, not all of our simulations detonate at a density high enough to generate substantial
amounts of nickel, we employ the mass-weighted average atomic mass number on the domain,
$\bar{A}$, as a more general metric describing the outcome of the simulation. A
final relevant metric is the total amount of energy $E$ released by the detonation.

Plotfiles of relevant quantities are created every 0.01 seconds.
The timesteps are limited to ensure that every multiple of this plotfile interval is
reached exactly, which means an effective maximum timestep of 0.01 seconds. Timesteps
are by default limited only by the hydrodynamic stability constraint and this plotfile
interval constraint, except where noted otherwise.

For a collision of two 0.64 $\msolar$ WDs with 128 zones in the radial dimension,
corresponding to a uniform spatial resolution of 312.5 km, the collision time is
$t_{\rm c} = 4.16\ \text{s}$, and the detonation time is $t_{\rm d} = 7.43\ \text{s}$.
The peak density at the detonation time is $2.3 \times 10^7\ \text{g\,/\,cm}^3$.
Approximately $0.21\ \msolar$ of nickel and $0.46\ \msolar$ of silicon is created,
corresponding to an $\bar{A}$ of about 29. Approximately $1.3 \times 10^{51}$
ergs of energy are released. Simulations with resolutions in this few hundred kilometer
range have been a common choice in earlier papers on white dwarf mergers and collisions
due to computational constraints; our goal is to see how higher resolution runs differ.

@@

At high resolution the collision time, $t_{\rm c} = 4.6\ \text{s}$, is later than the
collision time at low resolution, due to a sharper view of the contact point. However,
for the same reason, the detonation time is significantly earlier at $t_{\rm d} = 6.62\ \text{s}$,
as the peak density reaches the key density range of $10^7\ \text{g\,/\,cm}^3$ significantly earlier,
and this is what triggers the burning sufficient to generate an earlier detonation.
Approximately speaking, the later $t_{\rm d}$ is, the more
nickel is produced. As time passes, more white dwarf material slams into the center, compressing
and heating up the gas. Both effects are favorable to production of nickel. So, a later $t_{\rm d}$
implies that more material can be converted to nickel.
